---
title: "Lecture 22"
author: "Brecon Welch"
date: "2023-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Social Pressure
```{r}
library(tidyverse)
library(infer)
data(social, package = "qss")
social <- as_tibble(social)
social

social |> count(messages)

```



#Calculate ATE (sample difference in means)
```{r}
social |> 
  filter(messages %in% c("Control", "Neighbors")) |>
  group_by(messages) |> 
  summarize(turnout_prop = mean(primary2006)) |>
  pivot_wider(names_from = messages, values_from = turnout_prop) |>
  mutate(ATE = Neighbors - Control)
```
when we take the difference between two percentages = percentage point difference NOT percentage difference
when we take the difference between two numbers relative to the other = percentage difference --> ex. ((A-B)/B) How much we go from B to A divided by B (gives us percentage difference)


```{r}
social <- social |>
  filter(messages %in% c("Neighbors", "Control")) 

social |>
  mutate(messages_permute = sample(messages)) |>
  select(primary2006, messages, messages_permute)


social <- social |>
  mutate(turnout = if_else(primary2006 == 1, "Voted", "Didn't Vote"))
##create an outcome that is more labelled
##outcome of interest = turnout var 

est_ate <- social |>
  specify(turnout ~ messages, success = "Voted") |>
  calculate(stat = "diff in props", order = c("Neighbors", "Control"))
```
#what we are doing here is reshuffling who is in treatment group and control group because there is no treatment effect/outcome difference so it doesn't matter who is in treatment or control



```{r}
null_dist <- social |>
  specify(turnout ~ messages, success = "Voted") |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in props", order = c("Neighbors", "Control"))
  ##generate data under this null hypothesize

#want diff in pop means to be 0, so null is equal to independent 

null_dist
  
```
#specify what you are interested in


```{r}
null_dist |>
  visualize()
```
difference in proportions between control and treatment groups, we get are between -.5 and .5 percentage points; if there is truly no treatment effect then we should get proportion close to this


```{r}
ate_pval <- null_dist |>
  get_p_value(obs_stat = est_ate, direction = "both")
ate_pval
```
#not actually 0, just approximately 0

```{r}
null_dist |>
  visualize() + 
  shade_p_value(obs_stat = est_ate, direction = "both")
```

#calculate ATE in means
```{r}
library(gov50data)
trains <- trains |>
  mutate(treatment = if_else(treatment == 1, "Treated", "Control"))

trains_ate <- trains |>
  specify(numberim.post ~ treatment) |>
  calculate(stat = "diff in means", order = c("Treated", "Control"))
trains_ate
```
#have a difference in means here not a difference in propprtion



```{r}
trains_null <- trains |>
  specify(numberim.post ~ treatment) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "diff in means", order = c("Treated", "Control"))
trains_null

trains_null |>
  visualize()
```
#number is between 1-5 insteasd of 0-1 don't have to specify success - vote and have to say diff in means instead of diff in proportion
